---
title: "Almond Yield Model, Group C"
output: html_document
---

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(knitr)
library(kableExtra)

source('yield_model.R')
source('profit_model.R')
options(knitr.table.format = "latex") 
```

```{r load_data, message = FALSE}
# Read in climate data
clim <- read_delim("./clim.txt", delim=" ")
```

# Compute Yield

```{r model_yield_table}
model_results <- yield_model(clim)

kable(model_results %>% 
        rename(Year = year, "Almond Yield Anomaly (ton/acre)" = almond_yield), 
      digits=4,
      caption = "Modeled Almond Yield Anomaly 1989-2010",
      format = "html",
      table.attr = "style='width:50%;'") %>% 
  kable_classic() %>% 
  kable_styling(bootstrap_options = c("striped", "hover"),
                position = "left"
                )
```

# Visualize Yield

```{r model_yield_plot, warning = FALSE}
ggplot(data = model_results, aes(x = year, y = almond_yield)) +
  geom_line() +
  geom_point() +
  theme_light() +
  labs(x = "Year",
       y = "Almond Yield Anomaly (ton/acre)",
       title = "Modeled Almond Yield Anomaly 1989-2010")
```

# Additional Exploration

```{r}
rain <- clim %>% 
  group_by(year) %>% 
  summarize(sum_precip = sum(precip))

rain_yield <- full_join(model_results, rain)

ggplot(data = rain, aes(x = year, y = sum_precip)) + 
  geom_line() +
  geom_point() +
  theme_light() 
  
```

```{r}
linear_model <- lm(data = rain_yield, almond_yield ~ sum_precip)

summary.lm(linear_model)


ggplot(data = rain_yield, aes(x = sum_precip, 
                              y = almond_yield)) +
    geom_smooth(method = "lm")
```

Our findings demonstrate four years of almond yields below the anomaly (1989, 1999, 2003, 2004). Noticeable spikes in almond yield occur in 1995, 2005 and 2008. There seems to be a general pattern regarding consecutive years of yields slightly above/below the anomaly with on year spikes occurring every few years. By plotting annual precipitation, we see a similar pattern in the data, with high rainfall generally correlating with high almond yields. The linear regression also supports the correlation.

# Compute Profit

```{r}
profit <- profit_model(yield_results = model_results, price =  10)
```

